---
title: "RRDF gotcha"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Setup

First load the package.
```{r, eval=TRUE}
library(rrdf)
library(rrdfancillary)
```

# Example on what not to do

Here is how I understood what happens when the exactly same triples are inserted - only one triple is stored.

```{r, eval=TRUE}
store3<- new.rdf(ontology=FALSE)

sparql.rdf( store3, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )

SPARQLinsert<- '
PREFIX xsd:  <http://www.w3.org/2001/XMLSchema#>
INSERT DATA
{ 
  <http://example.org/subject1>  <http://example.org/property1> "mytext"^^xsd:string .
  <http://example.org/subject1>  <http://example.org/property1> "mytext"^^xsd:string .
}
'

cat(SPARQLinsert,"\n")

update.rdf( store3, SPARQLinsert )

sparql.rdf( store3, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )

```

Now, of course, it is always better only to store the values once, when it is intended to store one copy.
However, as I thougth that only one trippel is stored, I was less carefull in some of the code.

Here is what Apache/Jena does when using the RRDF interface.
```{r, eval=TRUE}
store<- new.rdf(ontology=FALSE)
add.data.triple( store, subject="http://example.org/subject1", predicate="http://example.org/property1", data="mytext", lang="en" )
add.data.triple( store, subject="http://example.org/subject1", predicate="http://example.org/property1", data="mytext", type="string" )
sparql.rdf( store, "select ?s ?p ?o where {?s ?p ?o}"  )
sparql.rdf( store, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )
```
The same tripel appears twice!

Mixing `INSERT DATA` and RRDF add.data.triple gives same result - two triples.

```{r, eval=TRUE}
store4<- new.rdf(ontology=FALSE)

sparql.rdf( store4, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )

SPARQLinsert<- '
PREFIX xsd:  <http://www.w3.org/2001/XMLSchema#>
INSERT DATA
{ 
  <http://example.org/subject1>  <http://example.org/property1> "mytext"^^xsd:string .
}
'

cat(SPARQLinsert,"\n")

update.rdf( store4, SPARQLinsert )

sparql.rdf( store4, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )

add.data.triple( store3, subject="http://example.org/subject1", predicate="http://example.org/property1", data="mytext", type="string" )

sparql.rdf( store4, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )

```

