---
title: "Demograhics table from CSV file"
author: "PhuseSubTeamAnalysisResults@example.org"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demograhics table from CSV file}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---


## Prelimnaries
The present version uses Apache Fuseki to access CDISC in RDF from rdf.cdisc.org.

Here is how to make a local copy of the rdf.cdisc.org - with Bash Shell on Linux.

```
export CDISCRDF_HOME=~/projects/phrmwg-cdisc # change to suit your setup
mkdir -p ${CDISCRDF_HOME}  # only if directory does not exist 
cd ${CDISCRDF_HOME}

git clone https://github.com/phuse-org/rdf.cdisc.org.git

mkdir rdf-data-cube
cd rdf-data-cube
wget http://publishing-statistical-data.googlecode.com/svn/trunk/specs/src/main/vocab/cube.ttl
cd ..

mkdir tdb

export JENA_HOME=/opt/apache-jena-2.12.1
${JENA_HOME}/bin/tdbloader2 --loc tdb rdf.cdisc.org/*/*.{ttl,owl} rdf-data-cube/*.ttl 
```

Start fuseki pointing to the TDB store created - note the update option makes it possible to upload triples as, say, turtle files.
```
cd ${CDISCRDF_HOME}
export FUSEKI_HOME=/opt/jena-fuseki-1.1.1
${FUSEKI_HOME}/fuseki-server --pages=${FUSEKI_HOME}/pages --update --loc=tdb /cdisc
```

## Code


```{r}

library (rrdf)

library(rrdfqbcrnd0)


#------------  Declarations    ------------------------------------------------
endpoint<- "http://localhost:3030/cdisc/query"

cubeVersion <- "0-5-2"  # For pav:Version (dot, notdash) and output file name
# dataDir <- "C:/_Sandbox/r/semanticWeb/data/"  # Input and output files
dataDir <- "./"  # Input and output files
prefixSourceFile  <- paste0(dataDir,"CubePrefixes.csv")    

DMskeletonSourceFile <- paste0(dataDir,"skeletonSource.csv") 
DMdsdName <- "dsd-demog"

# Cube Observation data       
DMobsFileName <-"demog.AR.csv" # Source of obs. Filename needed for cube metadata
DMobsFile     <- paste0(dataDir,DMobsFileName)  # Full path to obs. file

# should maybe get that from some other data - eg. stored in the cube
DMrecode.list= list(
    "sex"=list( 'F'='sex-F','M'='sex-M', '_ALL_'='sex-_ALL_' ),
    "race"= list('WHITE'='race-WHITE',
                       'BLACK OR AFRICAN AMERICAN'='race-BLACK_OR_AFRICAN_AMERICAN',
                       'AMERICAN INDIAN OR ALASKA NATIVE'='race-AMERICAN_INDIAN_OR_ALASKA_NATIVE',
                       '_ALL_'='race-_ALL_' ),
    "trt01a"=list('Placebo'='trt01a-Placebo',
                  'Xanomeline High Dose'='trt01a-Xanomeline_High_Dose',
                  'Xanomeline Low Dose'='trt01a-Xanomeline_Low_Dose',
                  '_ALL_'='trt01a-_ALL_' ),
    "saffl"=list('Y'='saffl-Y',
                        'N'='saffl-N',
                        '_ALL_'='saffl-_ALL_'),
    "factor"=list('quantity'='factor-QUANTITY',
                         'proportion'='factor-PROPORTION', 
                         'AGE'='factor-AGE' ),
    "procedure"= list('count'='procedure-COUNT',
                      'countdistinct'='procedure-COUNTDISTINCT',
                      'percent'='procedure-PERCENT', 
                      'mean'='procedure-MEAN', 
                      'stdev'='procedure-STDDEV', 
                      'min'='procedure-MIN', 
                      'median'='procedure-MEDIAN', 
                      'max'='procedure-MAX' )                        
     );

  procedure2format= list("count"="int",
                         "countdistinct"="int",
                            'percent'='double', 
                            'mean'='double', 
                            'stdev'='double', 
                            'min'='double', 
                            'median'='double', 
                            'max'='double'                         
     );


# Output file : DC-Demog-R-Vn-n-(n).TTL . Also used in dcat:distribution
DMdataCubeFileName  <- paste0("DC-Demog-R-V-",cubeVersion,".TTL")   
DMdataCubeFile      <- paste0(dataDir,DMdataCubeFileName) # Full path to cube
  
storeDM = new.rdf(ontology=FALSE)  # Initialize

############
# Prefixes #
###############################################################################
# Construct vars. to hold the prefix namespaces, using format:prefixPREFIXNAME.
# Columns in source file: prefix,namespace
# Vars. used  later in the construction of cube skeleton and cube obs.
#   Examples: prefixQB   holds value http://purl.org/linked-data/cube#
#             prefixRDFS holds value http://www.w3.org/2000/01/rdf-schema#
prefixSource = read.csv(prefixSourceFile) 
DMprefixlist= qb.def.prefixlist(storeDM, prefixSource)

############
# Skeleton #
###############################################################################
# Read the skeleton  specifications from CSV to dataframe
# Source in file: compType, compName, compLabel
# TODO: Replace some use of compLabel with a var : compNameClass, formed by
#     : Upcase of first letter of the compName value.
DMskeletonSource = read.csv(DMskeletonSourceFile) 

DMobsData = read.csv(DMobsFile) 
names(DMobsData)= tolower(names(DMobsData))

qb.buildSkeleton(storeDM, DMprefixlist, DMobsData, DMskeletonSource)

qb.buildDSD(storeDM, DMprefixlist, DMobsData, DMskeletonSource,
            dsdURIwoprefix="dataset-demog",
            dsdName=DMdsdName,
            extra=list(description="Cube with 6 Dimensions (factor, procedure, race, saffl, sex, trt01a), 2 Attributes (denominator, unit) and 1 measure (measure)",
comment="Example Demographics data supplied by Ian Fleming via R.. All dimensions have a Codelist and Range specified. Attributes applied from source data. Attributes as VALUES instead of URIs. Changes in current version:  codelists for factor, procedure, saffl, trt01a",
label="Demographics results data set.",
distribution=DMdataCubeFileName,
obsfilename=DMobsFileName,
title="Demographics Analysis Results"
 )            )

qb.buildObservations( store=storeDM, prefixlist=DMprefixlist, obsData=DMobsData, skeletonSource=DMskeletonSource, dsdURIwoprefix="dataset-demog", recode.list=DMrecode.list, procedure2format=procedure2format )

##########
# Output #
###############################################################################
DMoutcube = save.rdf(storeDM, filename=DMdataCubeFile, format="TURTLE")

###############################################################################
```


Now look at the cube. Note: by specifying prefix the output contains is shown using the prefixes.
Note for future: This may be a disadvantage if the value of the prefix, say ds, changes.

```{r, echo=FALSE, results='asis'}

checkDM = new.rdf(ontology=FALSE)  # Initialize
load.rdf("DC-Demog-R-V-0-5-2.TTL", format="TURTLE", appendTo= checkDM)
summarize.rdf(checkDM)


## ------------------------------------------------------------------------

myprefixes= list(
 "rdf"= "http://www.w3.org/1999/02/22-rdf-syntax-ns#" ,
 "skos"="http://www.w3.org/2004/02/skos/core#" ,
 "prov"="http://www.w3.org/ns/prov#" ,
 "rdfs"="http://www.w3.org/2000/01/rdf-schema#" ,
 "dcat"="http://www.w3.org/ns/dcat#" ,
 "owl"= "http://www.w3.org/2002/07/owl#" ,
 "xsd"= "http://www.w3.org/2001/XMLSchema#" ,
 "qb"=  "http://purl.org/linked-data/cube#" ,
 "pav"= "http://purl.org/pav" ,
 "dct"= "http://purl.org/dc/terms/" ,
 "mms"= "http://rdf.cdisc.org/mms#" ,
 "cts"= "http://rdf.cdisc.org/ct/schema#" ,
 "dccs"="http://www.example.org/dc/demog/dccs/" ,
 "code"="http://www.example.org/dc/code/" ,
 "ds"=  "http://www.example.org/dc/demog/dataset/" ,
 "prop"="http://www.example.org/dc/demog/prop/",
 "valid"="http://www.example.org/dc/demog/valid/"
)  ;


forsparqlprefix= paste("prefix", paste(names(myprefixes),":",sep=""), paste("<",myprefixes,">",sep=""),sep=" ",collapse=" ")

cube.observations1= sparql.rdf(checkDM,
   paste( forsparqlprefix,
     "select * where {?s ?p ?o .} limit 10"
     )
   );
knitr::kable(head(cube.observations1, 10))


cube.observations2= sparql.rdf(checkDM,
   paste( forsparqlprefix,
     "select * where { ?s a qb:Observation ; ?p ?o .}"
     )
   );
knitr::kable(head(cube.observations2, 10))
```
