---
title: "Demograhics table from CSV file"
author: "PhuseSubTeamAnalysisResults@example.org"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demograhics table from CSV file}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---


## Prelimnaries
The present version uses Apache Fuseki to access CDISC in RDF from rdf.cdisc.org.

Here is how to make a local copy of the rdf.cdisc.org - with Bash Shell on Linux.


```{r get-rdf-cdisc-org.sh, eval=FALSE, engine='bash'}
echo change below to suit your setup
export CDISCRDF_HOME=~/projects/phrmwg-cdisc 
echo only if directory does not exist 
mkdir -p ${CDISCRDF_HOME}  
cd ${CDISCRDF_HOME}

git clone https://github.com/phuse-org/rdf.cdisc.org.git

mkdir rdf-data-cube
cd rdf-data-cube
wget http://publishing-statistical-data.googlecode.com/svn/trunk/specs/src/main/vocab/cube.ttl
cd ..

mkdir tdb

export JENA_HOME=/opt/apache-jena-2.12.1
${JENA_HOME}/bin/tdbloader2 --loc tdb rdf.cdisc.org/*/*.{ttl,owl} rdf-data-cube/*.ttl 
```

Start fuseki pointing to the TDB store created - note the update option makes it possible to upload triples as, say, turtle files.


```{r start-fuseki.sh, eval=FALSE, engine='bash'}
cd ${CDISCRDF_HOME}
export FUSEKI_HOME=/opt/jena-fuseki-1.1.1
${FUSEKI_HOME}/fuseki-server --pages=${FUSEKI_HOME}/pages --update --loc=tdb /cdisc
```

### How to run from command prompt when developing

Start in the package root directory

Devtools should reply reply Loading rrdfqbcrnd0.

The knit command should generate .md file in the present work directory - the package root directory


```{r, eval=FALSE}
library(devtools)
devtools::load_all()

library(knitr)
knit("vignettes/dm-table-from-csv.Rmd")
```

Generate the pdf file using the following command
```{r makepdf.sh, eval=FALSE, engine='bash'}
pandoc dm-table-from-csv.md -o xx.pdf
```

The knitr::knit2html should generate a .html file in the present work directory - the package root directory

Generating HTML works nicely
```{r, eval=FALSE} 
knitr::knit2html("vignettes/dm-table-from-csv.Rmd")
```

The following does not work as the texi2dvi can not handle the dollar signs in the bash code above.
I suppose some other options have to be given. The work around is to use pandoc on the .md file generated as mentioned above.

```{r, eval=FALSE}
knitr::knit2pdf("vignettes/dm-table-from-csv.Rmd")
```

## Code


```{r}

library (rrdf)
library (xlsx)
library(rrdfqbcrnd0)


#------------  Declarations    ------------------------------------------------
endpoint<- "http://localhost:3030/cdisc/query"

# NOTE: For pav:Version (dot, notdash) and output file name
cubeVersion <- "0-5b"  # 0-5-3 using .XLSX for DM Skeleton  

# TODO: Differences between systems for tempdir() usage??
#TW dataDir <- "C:/_gitHub/rrdfqbcrnd0/inst/extdata/"  # TW requires this.  
dataDir <- tempdir()  # Input and output files: Used by MJA

generalprefixSourceFile<- system.file("extdata/sample-cfg", "CommonCubePrefixes.csv", package="rrdfqbcrnd0")
print(paste0("Reading from ", generalprefixSourceFile ))
if (! file.exists(generalprefixSourceFile) ) {
  # consider using try instead
  stop( paste0("Expected file ", generalprefixSourceFile, " does not exist" ))
  }
specificprefixSourceFile<- system.file("extdata/sample-cfg", "AECubePrefixes.csv", package="rrdfqbcrnd0")
if (! file.exists(specificprefixSourceFile) ) {
  # consider using try instead
  stop( paste0("Expected file ", specificprefixSourceFile, " does not exist" ))
  }

############
# Prefixes #
###############################################################################
# Construct list to hold the prefix namespaces, using format:prefixPREFIXNAME with key .
# Columns in source file: prefix,namespace
# Vars. used  later in the construction of cube skeleton and cube obs.
#   Examples: prefixQB   holds value http://purl.org/linked-data/cube#
#             prefixRDFS holds value http://www.w3.org/2000/01/rdf-schema#
generalprefixSource = read.csv(generalprefixSourceFile) 
specificprefixSource= read.csv(specificprefixSourceFile)
prefixSource= rbind(generalprefixSource, specificprefixSource)


#TW DMskeletonSourceFile <- system.file("extdata/sample-cfg", "skeletonSource-DM.csv", package="rrdfqbcrnd0")
# SWitching over from .csv to .XLSX
RDFCubeWorkbook = system.file("extdata/sample-cfg", "RDFCubeWorkBook.xlsx", package="rrdfqbcrnd0")

# Cube Observation data       
DMobsFile<- system.file("extdata/sample-cfg", "dm.AR.csv", package="rrdfqbcrnd0")


# should maybe get that from some other data - eg. stored in the cube
DMrecode.list= list(
    "sex"=list( 'F'='sex-F','M'='sex-M', '_ALL_'='sex-_ALL_' ),
    "race"= list('WHITE'='race-WHITE',
                       'BLACK OR AFRICAN AMERICAN'='race-BLACK_OR_AFRICAN_AMERICAN',
                       'AMERICAN INDIAN OR ALASKA NATIVE'='race-AMERICAN_INDIAN_OR_ALASKA_NATIVE',
                       '_ALL_'='race-_ALL_' ),
    "trt01a"=list('Placebo'='trt01a-Placebo',
                  'Xanomeline High Dose'='trt01a-Xanomeline_High_Dose',
                  'Xanomeline Low Dose'='trt01a-Xanomeline_Low_Dose',
                  '_ALL_'='trt01a-_ALL_' ),
    "saffl"=list('Y'='saffl-Y',
                        'N'='saffl-N',
                        '_ALL_'='saffl-_ALL_'),
    "factor"=list('quantity'='factor-QUANTITY',
                         'proportion'='factor-PROPORTION', 
                         'AGE'='factor-AGE' ),
    "procedure"= list('count'='procedure-COUNT',
                      'countdistinct'='procedure-COUNTDISTINCT',
                      'percent'='procedure-PERCENT', 
                      'mean'='procedure-MEAN', 
                      'stdev'='procedure-STDDEV', 
                      'min'='procedure-MIN', 
                      'median'='procedure-MEDIAN', 
                      'max'='procedure-MAX' )                        
     );

  procedure2format= list("count"="int",
                         "countdistinct"="int",
                            'percent'='double', 
                            'mean'='double', 
                            'stdev'='double', 
                            'min'='double', 
                            'median'='double', 
                            'max'='double'                         
     );


# Output file : DC-Demog-R-Vn-n-(n).TTL . Also used in dcat:distribution
DMdataCubeFileName  <- paste0("DC-Demog-R-V-",cubeVersion,".TTL")   
DMdataCubeFile      <- paste0(dataDir,DMdataCubeFileName) # Full path to cube
  
storeDM = new.rdf(ontology=FALSE)  # Initialize

# register prefixes
DMprefixlist= qb.def.prefixlist(storeDM, prefixSource)

############
# Skeleton #
###############################################################################
# Read the skeleton  specifications from CSV to dataframe
# Source in file: compType, compName, compLabel
# TODO: Replace some use of compLabel with a var : compNameClass, formed by
#     : Upcase of first letter of the compName value.
#TW DMskeletonSource = read.csv(DMskeletonSourceFile) 
#TODO: Convert "DM-Components" to parameter. Sheet name to match cube name "demog"?
DMCubeMetadata = read.xlsx(RDFCubeWorkbook,sheetName="DM-Components")
# Subset to the dimensions, attributes, and measure used to construct the skeleton
DMSkeletonSource <-DMCubeMetadata[grep("dimension|attribute|measure", DMCubeMetadata$compType),]

DMobsData = read.csv(DMobsFile) 
names(DMobsData)= tolower(names(DMobsData))

qb.buildSkeleton(storeDM, DMprefixlist, DMobsData, DMSkeletonSource)

#TODO: Remove the hard coding for "demog" in the statements below
qb.buildDSD(storeDM, DMprefixlist, DMobsData, DMSkeletonSource,
            dsdURIwoprefix="dataset-demog",
            dsdName="dsd-demog",
            extra=list(description="Cube with 6 Dimensions (factor, procedure, race, saffl, sex, trt01a), 2 Attributes (denominator, unit) and 1 measure (measure)",
comment="Example Demographics data supplied by Ian Fleming via R.. All dimensions have a Codelist and Range specified. Attributes applied from source data. Attributes as VALUES instead of URIs. Changes in current version:  codelists for factor, procedure, saffl, trt01a",
label="Demographics results data set.",
distribution=DMdataCubeFileName,
obsfilename=DMobsFile,
title="Demographics Analysis Results"
 ),
remote.endpoint=endpoint
            )

qb.buildObservations( store=storeDM, prefixlist=DMprefixlist, obsData=DMobsData, skeletonSource=DMSkeletonSource, dsdURIwoprefix="dataset-demog", recode.list=DMrecode.list, procedure2format=procedure2format )

##########
# Output #
###############################################################################
DMoutcube = save.rdf(storeDM, filename=DMdataCubeFile, format="TURTLE")

###############################################################################
```


Now look at the cube. Note: by specifying prefix the output contains is shown using the prefixes.
Note for future: This may be a disadvantage if the value of the prefix, say ds, changes.

```{r, echo=FALSE, results='asis'}

checkDM = new.rdf(ontology=FALSE)  # Initialize
load.rdf(DMdataCubeFile, format="TURTLE", appendTo= checkDM)
summarize.rdf(checkDM)


## ------------------------------------------------------------------------

myprefixes= list(
 "rdf"= "http://www.w3.org/1999/02/22-rdf-syntax-ns#" ,
 "skos"="http://www.w3.org/2004/02/skos/core#" ,
 "prov"="http://www.w3.org/ns/prov#" ,
 "rdfs"="http://www.w3.org/2000/01/rdf-schema#" ,
 "dcat"="http://www.w3.org/ns/dcat#" ,
 "owl"= "http://www.w3.org/2002/07/owl#" ,
 "xsd"= "http://www.w3.org/2001/XMLSchema#" ,
 "qb"=  "http://purl.org/linked-data/cube#" ,
 "pav"= "http://purl.org/pav" ,
 "dct"= "http://purl.org/dc/terms/" ,
 "mms"= "http://rdf.cdisc.org/mms#" ,
 "cts"= "http://rdf.cdisc.org/ct/schema#" ,
 "dccs"="http://www.example.org/dc/demog/dccs/" ,
 "code"="http://www.example.org/dc/code/" ,
 "ds"=  "http://www.example.org/dc/demog/dataset/" ,
 "prop"="http://www.example.org/dc/demog/prop/",
 "valid"="http://www.example.org/dc/demog/valid/"
)  ;


forsparqlprefix= paste("prefix", paste(names(myprefixes),":",sep=""), paste("<",myprefixes,">",sep=""),sep=" ",collapse=" ")

cube.observations1= sparql.rdf(checkDM,
   paste( forsparqlprefix,
     "select * where {?s ?p ?o .} limit 10"
     )
   );
knitr::kable(head(cube.observations1, 10))


cube.observations2= sparql.rdf(checkDM,
   paste( forsparqlprefix,
     "select * where { ?s a qb:Observation ; ?p ?o .}"
     )
   );
knitr::kable(head(cube.observations2, 10))
```
