---
title: "SPARQL scripts for RDF data cubes"
author: "PhuseSubTeamAnalysisResults@example.org"
date: "`r Sys.Date()`"
output: 
  html_document
  pdf_document
vignette: >
  %\VignetteIndexEntry{SPARQL scripts for RDF data cubes}
  %\VignetteEngine{knitr::knitr}
  %\usepackage[utf8]{inputenc}
  %\SweaveUTF8
---

# Setup

First load the package.
```{r, eval=TRUE}
library(rrdfqbcrnd0)
```

The display of SPARQL script in markdown is done by first creating a chunk, and then using the chunk with the highlight engine in knitr. The advantage of this approach is that all formatting is handled by external packages. To make the highlight output work in markdown two blanks has to be added at the end of line according to markdown syntax.
```{r helperfunction}
mdwrite<- function( sparqlStatements, refname ) {
fn<- file.path(tempdir(), paste0( refname, ".txt" ) )
cat( paste0("## @knitr ", refname), gsub("\\n", "  \n", sparql.rq), sep="  \n", file=fn)
knitr::read_chunk( fn, from=c(1))
invisible(fn)
}
```

```{r, eval=FALSE, echo=FALSE}
knitr::knit("vignettes/SPARQL-scripts-for-qb.Rmd")
markdown::markdownToHTML("SPARQL-scripts-for-qb.md",output="SPARQL-scripts-for-qb.html")

# makes very large file, and does not # at end of < >.
rmarkdown::render("vignettes/SPARQL-scripts-for-qb.Rmd", rmarkdown::html_document())
# the > at end of lines are shown as latex commands 
rmarkdown::render("vignettes/SPARQL-scripts-for-qb.Rmd", rmarkdown::pdf_document())

```

# SPARQL scripts

The rrdfqbcrnd0 package uses SPARQL scripts to access the RRDF triple
store.  This is implemented by having a function generating the SPARQL
script.  This vignettes shows the contents of the scripts. Some of the
scripts are parametrised by one or more paramteres. The parameter are
shown with as $p, $q etc, following the same convention as in the
(http://www.w3.org/TR/vocab-data-cube/#ic-20)[RDF Data Cube Vocabulary].

## SPARQL query for attributes in RDF data cube

```{r}
forsparqlprefix<- GetForSparqlPrefix( "$myQbName" )
sparql.rq<- GetAttributesSparqlQuery( forsparqlprefix )
mdwrite( sparql.rq, "sp1" )
```

```{r sp1, results='asis', engine='highlight', engine.opts='-S n3 --inline-css'}
```

## SPARQL query for observations from RDF data cube in workbook format

```{r}
forsparqlprefix<- GetForSparqlPrefix( "$myQbName" )
sparql.rq<- GetObservationsSparqlQuery( forsparqlprefix, "$myQbName", c("$dimA", "$dimB", "$dimC"), c("$attr1"))
mdwrite( sparql.rq, "sp2" )
```

```{r sp2, results='asis', engine='highlight', engine.opts='-S n3 --inline-css'}
```

# Generating this output
TODO(mja): enter the commands here as example in each of the .Rd files.
Then the output here can be generated from the .Rd files.

```{r, eval=FALSE, echo=FALSE}
db <- Rd_db("rrdfqbcrnd0")
keywords <- lapply(db, tools:::.Rd_get_metadata, "keyword")
db$qbIClist.Rd
tools:::RdTags(db$qbIClist.Rd)
titles <- lapply(db, tools:::.Rd_get_metadata, "title")
arguments <- lapply(db, tools:::.Rd_get_metadata, "arguments")
names <- lapply(db, tools:::.Rd_get_metadata, "name")
dofor<-grep("sparql",titles,ignore.case=TRUE) 
cbind(titles[dofor],arguments[dofor])
```
