---
title: "Two dimensional representation of RDF data cube"
author: "PhuseSubTeamAnalysisResults@example.org"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Two dimensional representation of RDF data cube}
  %\VignetteEngine{knitr::knitr}
  %\usepackage[utf8]{inputenc}
  %\SweaveUTF8
---

# Setup

First load the package.
```{r, eval=TRUE}
library(rrdfqbcrnd0)
```


# Current - 08-may-2015

The following code is under development. It creates HTML files
under extdata/sample-cfg showing a two dimensional representation
of the RDF data cube.

Here are some the features that are evaluated
- drag and drop of measure - could be used for creating a new table from existing
- store RDF as RDFa - could be used to embed the whole cube in the file; looks like it gets to big

Pending: including the rest of the cube into the
html. Consider if in each cell only the observation and measure should
be referenced and not all properties and objects for the observation.


## Using with RDFa

The function `MakeHTMLfromQb` is used to create the HTML file. When invoked with `useRDFa=TRUE` the generated HTML will contain RDFa markup.

The HTML includes green-turtle
(https://github.com/alexmilowski/green-turtle), jquery
(http://jquery.com/) and jqueryUI (http://jqueryui.com/).

In my setup I store the project under packages, and can symlink to the
files from the extdata/sample-cfg directory.


    cd extdata/sample-cfg
    ln -s ~/packages/green-turtle/build/RDFa.min.1.4.0.js 
    ln -s ~/packages/green-turtle/build/RDFaProcessor.min.1.4.0.js 
    ln -s ~/packages/jquery-2.1.3.min/jquery-2.1.3.min.js .
    ln -s ~/packages/jquery-ui-1.11.3.custom .




```{r, echo=TRUE}
MakeTable<- function( dataCubeFile, htmlfile, rowdim, coldim, idrow, idcol ) {
    store <- new.rdf()  # Initialize
    cat("Loading ", dataCubeFile, "\n")
    load.rdf(dataCubeFile, format="TURTLE", appendTo= store)
    summarize.rdf(store)
    dsdName<- GetDsdNameFromCube( store )
    domainName<- GetDomainNameFromCube( store )
    forsparqlprefix<- GetForSparqlPrefix( domainName )

    dimensions<- sparql.rdf(store, GetDimensionsSparqlQuery( forsparqlprefix ) )
    attributesDf<- sparql.rdf(store, GetAttributesSparqlQuery( forsparqlprefix ))

    outhtmlfile<- MakeHTMLfromQb( store, forsparqlprefix, dsdName, domainName,
                                 dimensions, rowdim, coldim, idrow, idcol,
                                 htmlfile, useRDFa=FALSE, compactDimColumns=FALSE)

    outhtmlfile
}

```

## Create AE html example

```{r, echo=TRUE}
dataCubeFile<- system.file("extdata/sample-rdf", "DC-AE-sample.ttl", package="rrdfqbcrnd0")
htmlfile<- file.path(system.file("extdata/sample-cfg", package="rrdfqbcrnd0"), "DC-AE-sample.html")
rowdim<- c("crnd-attribute:rowno", "crnd-dimension:aesoc", "crnd-dimension:aedecod" )
coldim<- c("crnd-attribute:colno", "crnd-attribute:cellpartno", "crnd-dimension:trta", "crnd-dimension:factor", "crnd-dimension:procedure" )
# idrow is a function of rowdim; writing it directly is easier for now
idrow<-  c( "aesocvalue", "aedecodvalue" )
idcol<-  c( "crnd-dimension:trta" )
resHtmlFile<- MakeTable( dataCubeFile, htmlfile, rowdim, coldim, idrow, idcol )
cat("HTML stored as: ", normalizePath(resHtmlFile), "\n")
```

## Create DEMO html example

```{r, echo=TRUE}
dataCubeFile<- system.file("extdata/sample-rdf", "DC-DEMO-sample.ttl", package="rrdfqbcrnd0")
htmlfile<- file.path(system.file("extdata/sample-cfg", package="rrdfqbcrnd0"), "DC-DEMO-sample.html")

coldim<- c("crnd-attribute:colno", "crnd-attribute:cellpartno", "crnd-dimension:trt01a", "crnd-dimension:factor", "crnd-dimension:procedure" )
rowdim<- c("crnd-attribute:rowno", "crnd-dimension:agegr1", "crnd-dimension:race", "crnd-dimension:ethnic", "crnd-dimension:sex" )
idrow<-  c( "agegr1value", "racevalue", "ethnicvalue", "sexvalue" )
idcol<-  c( "crnd-dimension:trt01a" )

resHtmlFile<- MakeTable( dataCubeFile, htmlfile, rowdim, coldim, idrow, idcol )
cat("HTML stored as: ", normalizePath(resHtmlFile), "\n")

```

## For development

### Setup

```{r, eval=TRUE}
store <- new.rdf()  # Initialize
cat("Loading ", dataCubeFile, "\n")
load.rdf(dataCubeFile, format="TURTLE", appendTo= store)
summarize.rdf(store)

dsdName<- GetDsdNameFromCube( store )
domainName<- GetDomainNameFromCube( store )
forsparqlprefix<- GetForSparqlPrefix( domainName )
```

### Check GetTwoDimTableFromQb

```{r, eval=TRUE}
dataCubeFile<- system.file("extdata/sample-rdf", "DC-DEMO-sample.ttl", package="rrdfqbcrnd0")
store <- new.rdf()  # Initialize
cat("Loading ", dataCubeFile, "\n")
load.rdf(dataCubeFile, format="TURTLE", appendTo= store)
dsdName<- GetDsdNameFromCube( store )
domainName<- GetDomainNameFromCube( store )
forsparqlprefix<- GetForSparqlPrefix( domainName )
coldim<- c("crnd-attribute:colno", "crnd-attribute:cellpartno", "crnd-dimension:trt01a" )
rowdim<- c("crnd-attribute:rowno", "crnd-dimension:agegr1", "crnd-dimension:race",
           "crnd-dimension:ethnic", "crnd-dimension:sex", "crnd-dimension:procedure" )
xqbtest<- GetTwoDimTableFromQb( store, forsparqlprefix, domainName, rowdim, coldim )

dimensionsRq <- GetDimensionsSparqlQuery( forsparqlprefix )
dimensions<- sparql.rdf(store, dimensionsRq)

attributesRq<- GetAttributesSparqlQuery( forsparqlprefix )
attributes<- sparql.rdf(store, attributesRq)

observationsDescriptionRq<- GetObservationsWithDescriptionSparqlQuery( forsparqlprefix, domainName, dimensions, attributes )
observationsDesc<- as.data.frame(sparql.rdf(store, observationsDescriptionRq ), stringsAsFactors=FALSE)


```

### View a specific observation.

The observation is specified in the values part of the SPARQL query.

```{r, eval=TRUE}

cube.observations.rq<-  paste( forsparqlprefix,
'
select *
where { 
?s a qb:Observation ; 
?p ?o .
values (?s) {
(ds:obs029)
}
}
',
"\n"                               
)

cube.observations<- sparql.rdf(store, cube.observations.rq)
knitr::kable(cube.observations)

```

### View formating of measure

```{r, eval=TRUE}
cube.measurefmt.rq<-  paste( forsparqlprefix,
'
select distinct ?procedure ?measurefmt
where { 
?s a qb:Observation ; 
crnd-dimension:procedure ?procedure ;
crnd-attribute:measurefmt ?measurefmt .
}
',
"\n"                               
)

cube.measurefmt<- sparql.rdf(store, cube.measurefmt.rq)
knitr::kable(cube.measurefmt)

```

### Columns

The contents of the 
```{r, echo=TRUE}
cols.rq<- GetRownoColnoCellpartnoSparqlQuery( forsparqlprefix )
cols<- data.frame(sparql.rdf(store, cols.rq))
knitr::kable(cols)

```


# Session information
```{r, echo=TRUE}
sessionInfo()
```

