---
title: "Two dimensional representation of RDF data cube"
author: "PhuseSubTeamAnalysisResults@example.org"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Two dimensional representation of RDF data cube}
  %\VignetteEngine{knitr::knitr}
  %\usepackage[utf8]{inputenc}
  %\SweaveUTF8
---

# Setup

First load the package.
```{r, eval=TRUE}
library(rrdfqbcrnd0)
```

# Loading DEMO data 

The DEMO data exists as a turtle file in the sample-rdf directory.

```{r, eval=TRUE}

dataCubeFile<- system.file("extdata/sample-rdf", "DC-DEMO-sample.TTL", package="rrdfqbcrnd0")
store <- new.rdf()  # Initialize
cat("Loading ", dataCubeFile, "\n")
load.rdf(dataCubeFile, format="TURTLE", appendTo= store)
summarize.rdf(store)

dsdName<- GetDsdNameFromCube( store )
domainName<- GetDomainNameFromCube( store )
forsparqlprefix<- GetForSparqlPrefix( domainName )
```

View a specific observation. The observation is specified in the values part of the SPARQL query.

```{r, eval=TRUE}
cube.observations.rq<-  paste( forsparqlprefix,
'
select *
where { 
?s a qb:Observation ; 
?p ?o .
values (?s) {
(ds:obs29)
}
}
',
"\n"                               
)

cube.observations<- sparql.rdf(store, cube.observations.rq)
knitr::kable(cube.observations)

```


# Current - 15-mar-2015

The following code is under development. It creates a HTML file
extdata/sample-cfg/test.html showing a two dimensional representation
of the table with drag and drop. The observations are stored as RDFa
in the HTML. Pending: including the rest of the cube into the
html. Consider if in each cell only the observation and measure should
be referenced and not all properties and objects for the observation.


The HTML includes green-turtle
(https://github.com/alexmilowski/green-turtle), jquery
(http://jquery.com/) and jqueryUI (http://jqueryui.com/).  In my setup
I store the project under packages, and can symlink to the files from the
extdata/sample-cfg directory.

    cd extdata/sample-cfg
    ln -s ~/packages/green-turtle/build/RDFa.min.1.4.0.js 
    ln -s ~/packages/green-turtle/build/RDFaProcessor.min.1.4.0.js 
    ln -s ~/packages/jquery-2.1.3.min/jquery-2.1.3.min.js .
    ln -s ~/packages/jquery-ui-1.11.3.custom .

```{r, echo=TRUE}
cols.rq<-  paste( forsparqlprefix,
'
select distinct ?colno ?cellpartno
where { 
?s a qb:Observation ; 
crnd-attribute:colno ?colno ;
crnd-attribute:cellpartno ?cellpartno .
}
order by ?colno ?cellpartno
',
"\n"                               
)

cols<- sparql.rdf(store, cols.rq)
knitr::kable(cols)

dimensions<- sparql.rdf(store, GetDimensionsSparqlQuery( forsparqlprefix ) )
attributesDf<- sparql.rdf(store, GetAttributesSparqlQuery( forsparqlprefix ))

rowdim<- c("crnd-attribute:rowno", "crnd-dimension:agegr1", "crnd-dimension:race", "crnd-dimension:ethnic", "crnd-dimension:sex" )
coldim<- c("crnd-attribute:colno", "crnd-attribute:cellpartno", "crnd-dimension:trt01a" )

qbtest<- GetTwoDimTableFromQb( store, forsparqlprefix, domainName, rowdim, coldim )

## names(attributes(qbtest))
## options(width=200)
## knitr::kable(qbtest[order(strtoi(qbtest$rowno)),])

oDx<-attr(qbtest,"observationsDesc")
oDxx<- oDx[! is.na(oDx$s),]
oD<- oDxx[order(strtoi(oDxx$rowno)),]
presrowvarindex<- unique(oD$rowno)
colvarindex<- unique(oD$colno)
cellpartnoindex<- unique(oD$cellpartno)
presrowvarindex
colvarindex
cellpartnoindex
oD[,c("s","rowno","colno","cellpartno")]

presrowvarvalue<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:(.*)","\\1value",rowdim)
presrowvarIRI<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:(.*)","\\1IRI",rowdim)
presrowvarlabel<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:(.*)","\\1label",rowdim)

## add code for embedding the cube as turtle
## determine cube compontents except observations,
## as the observations are stored as RDFa
file<- file.path(system.file("extdata/sample-cfg", package="rrdfqbcrnd0"), "test.html")
# file<- file.path(tempdir(),"test.html")
append<- TRUE
print(file)
cat("<!DOCTYPE HTML>\n", file=file, append=FALSE)
cat('
<html>
  <head>
<meta charset="UTF-8">
<title>DEMO table as html</title>
   <script src="jquery-2.1.3.min.js"></script>
   <link rel="stylesheet" href="jquery-ui-1.11.3.custom/jquery-ui.css"/>
   <script src="jquery-ui-1.11.3.custom/jquery-ui.min.js"></script>
   <script src="RDFa.min.1.4.0.js"></script>

<style>
#table {
    line-height:30px;
    background-color:#eeeeee;
    height:1000px;
    width:750px;
    float:left;
    padding:5px;
}
#drop{
    width:300px;
    background-color:green;
    float:left;
    padding:10px;
}
</style>

</head>
<script>
"use strict";

function allowDrop(ev)
{
ev.preventDefault();
}

function drag(ev)
{
ev.dataTransfer.setData("Text",ev.target.id);
console.log("Dragging: ", ev.target.id);
}

function drop(ev)
{
ev.preventDefault();
var data=ev.dataTransfer.getData("Text");
console.log("Dropping: ", data);
// from http://stackoverflow.com/questions/13007582/html5-drag-and-copy
var nodeCopy = document.getElementById(data).cloneNode(true);
nodeCopy.id = "copy"+nodeCopy.id;
// end from http://stackoverflow.com/questions/13007582/html5-drag-and-copy
var newelem = document.createElement("P");
newelem.appendChild(nodeCopy);
ev.target.appendChild(newelem);
}

$(document).ready(function(){

GreenTurtle.attach(document)

})
                  
</script>
<body>
<h1>DEMO table as RDFa</h1>
'
, file=file, append=TRUE)

cat('<div id="container">', file=file, append=TRUE)

cat("<div id='table'>\n", file=file, append=TRUE)
cat("<table border>\n", file=file, append=TRUE)
or<-1

## make the header row(s) for the columns

for (rr in presrowvarindex) {
  cat("<tr>", file=file, append=TRUE)
print(rr)
  ## make the row identification
  for (cc in colvarindex) {
print(cc)
    cat("<td>", file=file, append=TRUE)
cpindex<-0
    for (cp in cellpartnoindex) {
print(cp)
cpindex<- cpindex+1
if (cpindex>1) {
## separator between cells should be taken from data
      cat(" ", file=file, append=TRUE)
}
    if (oD$rowno[or]==rr & oD$colno[or]==cc & oD$cellpartno[or]==cp ) {
## The observation
      ## next line is for simple fly-over
      cat(paste0("<a title=\"", oD$measureIRI[or], "\">"), file=file, append=TRUE)

      cat(paste0('<span ', 'id="', gsub("ds:","",oD$s[or]), '"',
                 'resource="', oD$s[or],'"',
                 ' typeof="qb:Observation" draggable="true" ondragstart="drag(event)" >'),
          file=file, append=TRUE)
cat(paste0('<span property="qb:dataSet" resource="', 'ds:', dsdName,'">'), file=file, append=TRUE)
      
for (prop in dimensions) {
cat( paste0('<span property="', prop, '"', ' resource="', oD[or, gsub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", prop)], '">'), file=file, append=TRUE)
}

## formatting to applied to measure
      cat(paste0(oD$measure[or]), file=file, append=TRUE)

for (prop in dimensions) {
cat( '</span>', file=file, append=TRUE)
}
cat( '</span>', file=file, append=TRUE)
cat( '</span>', file=file, append=TRUE)

      cat(paste0("</a>"), file=file, append=TRUE)
      or<- or+1
    }
    }
    cat("</td>", file=file, append=TRUE)
  }
cat("</tr>", file=file, append=TRUE)
  }
cat("</table>\n", file=file, append=TRUE)
cat("</div>\n", file=file, append=TRUE)

cat('
<div id="droparea">
Drag and drop over the green text below.
<table>
<tr><td>
<span  style="width:100px" id="drop" ondrop="drop(event)" ondragover="allowDrop(event)">Drop here...</span>
</td></tr>
</table>
</div> 
',  file=file, append=TRUE)                 
cat('
</div>
</body>
</html>
', file=file, append=TRUE)
print(file)

```

# Session information
```{r, echo=TRUE}
sessionInfo()
```

