---
title: "POC: Clinical Research And Development (CRND) RDF Data Cube Structure created Using Literate Programming Techniques"
author: "Marc Andersen"
date: "`r Sys.Date()`"
output:
  knitr:::html_vignette:
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{POC: Clinical Research And Development (CRND) RDF Data Cube Structure created Using Literate Programming Techniques}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

# Thanks to

[Literate Programming](http://en.wikipedia.org/wiki/Literate_programming) (LP) technique

Setup for the Literate Programming based on the **rlp** package is available on Github: https://github.com/yihui/rlp. 

# Setup

Following the **rlp** package the **knitr** chunk options are set as:

```{r setup, purl=FALSE}
knitr::opts_chunk$set(eval = FALSE, tidy = FALSE)
```
The R code:
* is written in this document
* extracted using the `purl` command defined in the Makefile in the package root directory.
* documented using **roxygen2** syntax.

The `Makefile` is

```{r engine='cat', engine.opts=list(lang='Makefile'), code=readLines('../Makefile'), purl=FALSE}
```

To execute R-code but not put it in the extracted output, use: `{r eval=TRUE, purl=FALSE}`.


```{r}
##' Dummy function used for defining all paramter with Roxygen2
##' 
##' List all parameters used
##' @param store the RRDF store
##' @param defDataSet defines the cube dataset
##' @param defDSD define the cube data set definition. data.frame with one row for each item
##' @param defCodeList define the cube codelist. data.frame with one row for each code level, and columns for dimension name (variable name), value (as charactere), RDF data type (?)
##' @return not defined
dummy<- function(store,defDataSet,defDSD,defCodeList) {
FALSE
}
```

The expansion of the URIs should be handled using the Jena function expandPrefix.
See http://stackoverflow.com/questions/5698583/resolve-prefix-programatically-jena

```{r}
##' Helper function used for adding triples stored in a data.frame
##' 
##' It is assumed that the first 4 columns contains the subject, property and object and datatype
##' The subject and property are expanded URI. The object is character.
##' @inheritParams addTriplesFromDF
##' ##' @return not defined
expandTriplesFromDF<- function(RPrefixMap,triplesDF) {
    expandedTriplesDF<- triplesDF
}
```

```{r}
##' Helper function used for adding triples stored in a data.frame
##' 
##' It is assumed that the first 4 columns contains the subject, property and object and datatype
##' The subject and property are expanded URI. The object is character.
##' @inheritParams dummy
##' @param triplesDF
##' @return not defined
addTriplesFromDF<- function(store,triplesDF) {
FALSE
}
```

