---
title: "Create simple RDF data qube"
author: "PhuseSubTeamAnalysisResults@example.org"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create simple RDF data qube}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

## Introduction
This vignette shows how to create a very simple RDF data cube from 
- data 
- metadata  

## Setup
```{r}
library (rrdf)
library(rrdfqbcrnd0)
```

## Define data
```{r}
obsData<- data.frame(
  category=c("AA-group", "BB-group"),
  procedure=c("count", "count" ),
  factor=c("quantity", "quantity" ),
  unit=c("subject", "subject" ),
  denominator=c(" ", " "),
  measure=c( 123, 456 ),
  stringsAsFactors=FALSE  )
knitr::kable(obsData)
```

## Define meta data
```{r}
cubeMetadata<- data.frame(
  compType=c("dimension", "dimension", "dimension", "unit", "denominator", "measure", "metadata"),
  compName=c("category", "procedure", "factor", "attribute", "attribute", "measure", "domainName"),
  codeType=c("DATA", "DATA", "DATA", " ", " ", "<NA>","<NA>"),
  nciDomainValue=c(" "," "," "," ", " ", " "," "),
  compLabel=c("Category", "Statistical procedure", "Type of procedure", "Result", "Unit", "Denominator", "EXAMPLE"),
  Comment=c(" "," "," "," "," "," "," "),
  stringsAsFactors=FALSE  )
knitr::kable(cubeMetadata)
```

## Create RDF data cube

The RDF data cube for the data above is created using
```{r}
outcube<- BuildCubeFromDataFrames(cubeMetadata, obsData )
```

The RDF data cube is serialized in turtle format and stored as
```{r}
outcube
```

## Looking at the RDF data cube

Now look at the generated cubes by loading the turle files. Note: by specifying prefix the output contains is shown using the prefixes.
Note for future: This may be a disadvantage if the value of the prefix, say ds, changes.

```{r}

dataCubeFile<- outcube
```

```{r}
# the rest of the code only depends on the value of dataCubeFile 
checkCube <- new.rdf(ontology=FALSE)  # Initialize
load.rdf(dataCubeFile, format="TURTLE", appendTo= checkCube)
summarize.rdf(checkCube)
# determine the domain name; used for defining prefixes
# TODO: reconsider the use of domain specific prefixes

# TODO: make this simpler - the only purpose is find the dsdName
# TODO: a qb:DataStructureDefinition, and for domainname, say, the DM in ds:dsd-DM
tempstr<- as.character(sparql.rdf(checkCube, "select ?s where { ?s a <http://purl.org/linked-data/cube#DataStructureDefinition> } limit 1"))
tempstrvec<- unlist(strsplit( tempstr, "/"))
dsdName<- tempstrvec[length(tempstrvec)]
domainName<- strsplit(dsdName,"-")[[1]][[2]]

common.prefixes <- data.frame(
    prefix=gsub("^prefix","",names(Get.default.crnd.prefixes())),
    namespace=as.character(Get.default.crnd.prefixes() )
    )
custom.prefixes <-Get.qb.crnd.prefixes(tolower(domainName))

forsparqlprefix<- Get.rq.prefix.df(rbind(common.prefixes, custom.prefixes))
```

The next statement shows the first 10 triples in the cube.
```{r, echo=FALSE}

cube.observations1.rq<- paste( forsparqlprefix,
'
select *
where {?s ?p ?o .}
limit 10
',
"\n"
)

cube.observations1<- sparql.rdf(checkCube, cube.observations1.rq  )
knitr::kable(head(cube.observations1, 10))
```

The next statement shows the first 30 triples in the cube, where the subject is a qb:Observation.
```{r}

cube.observations2.rq<-  paste( forsparqlprefix,
'
select *
where { ?s a qb:Observation ; ?p ?o .}
limit 30
',
"\n"                               
)

cube.observations2<- sparql.rdf(checkCube, cube.observations2.rq)
knitr::kable(head(cube.observations2, 10))

```

The codelists are shown in the next output.
```{r, echo=FALSE}

codelists.rq<-   paste(forsparqlprefix,
'
select distinct ?p ?cl ?prefLabel
where {
?DataStructureDefinition a qb:DataStructureDefinition ;
   qb:component ?component .
?component a qb:ComponentSpecification .
?component qb:dimension ?p .
?p qb:codeList ?c .
?c skos:hasTopConcept ?cl .
?cl skos:prefLabel ?prefLabel .
values ( ?DataStructureDefinition ) {
',
paste0( "(", "ds:", dsdName, ")"),
'
}
}
order by ?p ?cl ?prefLabel
'
)

cube.codelists<- as.data.frame(sparql.rdf(checkCube, codelists.rq), stringsAsFactors=FALSE)
# TODO instead of gsub make a more straightforward way
# TOTO this involves a new version of the ph.recode function
cube.codelists$vn<- gsub("prop:","",cube.codelists$p)
cube.codelists$clc<- gsub("code:","",cube.codelists$cl)
knitr::kable(print(cube.codelists[,c("vn", "clc", "prefLabel")]))

```

The dimensions are shown in the next output.
```{r}

cube.dimensions.rq<- paste(forsparqlprefix,
'
select * where
{ [] qb:dimension ?p .  }
',
"\n"
)
cube.dimensions<- as.data.frame(sparql.rdf(checkCube, cube.dimensions.rq), stringsAsFactors=FALSE)
knitr::kable(print(cube.dimensions))

```

And finally the SPARQL query for observations.
```{r, echo=FALSE}

cube.dimensionsattr<- sparql.rdf(checkCube,
  paste(forsparqlprefix,
"select * where { {[] qb:dimension ?p . } union {  ?p a qb:AttributeProperty . } }"
))

cube.observations.rq<-  paste( forsparqlprefix,
    "select * where {",
    "?s a qb:Observation  ;", "\n",
    paste("       qb:dataSet",  paste0( "ds:", "dataset", "-", domainName), " ;", sep=" ", collapse="\n"), "\n",
    paste0( cube.dimensionsattr, " ", sub("prop:", "?", cube.dimensionsattr), ";", collapse="\n"),
    "\n",
    "prop:measure      ?measure ;      \n",
    paste0( "optional{ ", sub("prop:", "?", cube.dimensionsattr), " ",
           "skos:prefLabel",
           " ",
           sub("prop:", "?", cube.dimensionsattr), "value" ,
           " . ", "}",
           collapse="\n"),
    "\n",
    "} ",
    "\n"
   )

```

This is the query for getting the observations
```{r}

cat(cube.observations.rq)
```

And finally the observations.
```{r}


cube.observations<- as.data.frame(sparql.rdf(checkCube, cube.observations.rq ), stringsAsFactors=FALSE)
knitr::kable(cube.observations[,c(paste0(sub("prop:", "", cube.dimensionsattr), "value"),"measure")])


```
