---
title: "RDF data cube SPARQL queries"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: 
  %\VignetteIndexEntry{RDF data cube SPARQL queries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Setup
```{r, results='asis', eval=TRUE}
devtools::load_all(pkg="..")

#library(rrdf)

```

# Load example cube and define SPARQL prefixes

```{r, results='asis', eval=TRUE}

qbfile <- file.path(system.file("extdata/sample-rdf", "example.ttl", package="rrdfqb") )
qbfile <- file.path(system.file("extdata/sample-rdf", "DC-DEMO-sample.ttl", package="rrdfqb") )
cube <- new.rdf(ontology=FALSE)
invisible(load.rdf( qbfile, format="TURTLE", cube))

SPARQLprefix<- '
prefix rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>  
prefix skos: <http://www.w3.org/2004/02/skos/core#>  
prefix prov: <http://www.w3.org/ns/prov#>  
prefix rdfs: <http://www.w3.org/2000/01/rdf-schema#>  
prefix dcat: <http://www.w3.org/ns/dcat#>  
prefix owl: <http://www.w3.org/2002/07/owl#>  
prefix xsd: <http://www.w3.org/2001/XMLSchema#>  
prefix qb: <http://purl.org/linked-data/cube#>  
prefix pav: <http://purl.org/pav>  
prefix dct: <http://purl.org/dc/terms/>  
'
```

# Get data set

```{r, results='asis', eval=TRUE}

SPARQLquery<- '
select ?dataset
where {
?dataset a qb:DataSet .
} 
'

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
result <- sparql.rdf(cube, SPARQLscript )
knitr::kable(result)

```

# Get the data structure definition in the store

```{r, results='asis', eval=TRUE}

SPARQLquery<- '
select ?dataset ?dsd
where {
?dataset a qb:DataSet ;
  qb:structure ?dsd .
 ?dsd a qb:DataStructureDefinition  .
} 
'

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
result <- sparql.rdf(cube, SPARQLscript )
knitr::kable(result)

```

# Get the componentes in the data structure definition in the store

```{r, results='asis', eval=TRUE}

SPARQLquery<- '
select ?dataset ?dsd ?component ?componentPropery ?componentReference 
where {
?dataset a qb:DataSet ;
  qb:structure ?dsd .
?dsd a qb:DataStructureDefinition;
qb:component ?component .
?component ?componentPropery ?componentReference .
} 
'

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
result <- sparql.rdf(cube, SPARQLscript )
knitr::kable(result)

```

# Get the dimension, measures and attributes

```{r, results='asis', eval=TRUE}
SPARQLquery<- '
select ?dataset ?component ?isblankcomponent ?componentProperty ?componentReference ?ordercp ?order ?label
where {
?dataset a qb:DataSet ;
  qb:structure ?dsd .
?dsd a qb:DataStructureDefinition;
qb:component ?component .
?component ?componentProperty ?componentReference .
optional{ ?component qb:order ?temporder . }
BIND( ISBLANK(?component) as ?isblankcomponent )
optional{ ?componentReference rdfs:label ?templabel . }
BIND(
if( ?componentProperty = qb:dimension, 1001, 
   if( ?componentProperty = qb:measure, 2001,
      if( ?componentProperty = qb:attribute, 3001, 9001 )
        )
  )
 as ?ordercp )
BIND( COALESCE( ?temporder, 0 )  as ?order )
BIND( COALESCE( ?templabel, "" )  as ?label )
values ( ?componentProperty ) {
(qb:attribute)
(qb:dimension)
(qb:measure)
}
}
order by ?dataset ?ordercp ?order ?componentReference
'
SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
result <- data.frame(sparql.rdf(cube, SPARQLscript ),stringsAsFactors=FALSE)
knitr::kable(result)
```

Add CONSTRUCT query that defines the cube from the SPARQL results above.

# Make SPARQL query for getting observations for example.ttl

This will not work as the cube is not normalized.


```{r, results='asis', eval=FALSE}
genfor<- result[result$dataset==result$dataset[1],]
genfor$colnames<- sapply(strsplit(genfor$componentReference, "#"), "[", 2)

SPARQLquery<- paste( "select * where {",
      "?obs a qb:Observation;",
      "# qb:dataSet  ?dataset ; ",
      paste("",
          "<", genfor$componentReference, ">", " ", "?", genfor$colnames, " ;", sep="", collapse="\n"
      ),
      ".", 
      paste( "#values (?dataset) { (",
            "<", result$dataset[1], ">",
            ") }", sep="", collapse="\n"
            ),
      "}", sep="\n", collapse="\n"
      )

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
cat(SPARQLscript)
oresult <- data.frame(sparql.rdf(cube, SPARQLscript ),stringsAsFactors=FALSE)
knitr::kable(oresult)



```

# Make SPARQL query for getting observations for example.ttl

This will not work as the cube is not normalized.


```{r, results='asis', eval=FALSE}
genfor<- result[result$dataset==result$dataset[1],]
genfor$colnames<- sapply(strsplit(genfor$componentReference, "#"), "[", 2)

SPARQLquery<- paste( "select * where {",
      "?obs a qb:Observation;",
      "# qb:dataSet  ?dataset ; ",
      paste("",
          "<", genfor$componentReference, ">", " ", "?", genfor$colnames, " ;", sep="", collapse="\n"
      ),
      ".", 
      paste( "#values (?dataset) { (",
            "<", result$dataset[1], ">",
            ") }", sep="", collapse="\n"
            ),
      "}", sep="\n", collapse="\n"
      )

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
cat(SPARQLscript)
oresult <- data.frame(sparql.rdf(cube, SPARQLscript ),stringsAsFactors=FALSE)
knitr::kable(oresult)

qq<-'
prefix rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>  
prefix skos: <http://www.w3.org/2004/02/skos/core#>  
prefix prov: <http://www.w3.org/ns/prov#>  
prefix rdfs: <http://www.w3.org/2000/01/rdf-schema#>  
prefix dcat: <http://www.w3.org/ns/dcat#>  
prefix owl: <http://www.w3.org/2002/07/owl#>  
prefix xsd: <http://www.w3.org/2001/XMLSchema#>  
prefix qb: <http://purl.org/linked-data/cube#>  
prefix pav: <http://purl.org/pav>  
prefix dct: <http://purl.org/dc/terms/>  
 select * where {
?obs a qb:Observation;
 qb:dataSet  ?dataset ; 
 <http://example.org/ns#refArea> ?refArea ;
 <http://example.org/ns#refPeriod> ?refPeriod ;
 <http://purl.org/linked-data/sdmx/2009/dimension#sex> ?sex ;
 <http://example.org/ns#lifeExpectancy> ?lifeExpectancy ;
# <http://purl.org/linked-data/sdmx/2009/attribute#unitMeasure> ?unitMeasure ;
.
#values (?dataset) { (<http://example.org/ns#dataset-le3>) }
}
'
oresult <- data.frame(sparql.rdf(cube, qq ),stringsAsFactors=FALSE)
knitr::kable(oresult)


```
