---
title: "Create prefixes for use with CDISC standards in RDF"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create prefixes for use with CDISC standards in RDF}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

# Prefixes for the package

This script creates a list with the prefix used for the package.
The prefix corresponds to the usual values.

```{r, results='asis', eval=TRUE}

library(devtools)

qbCDISCprefixes<-  list(
 "rdf"= "http://www.w3.org/1999/02/22-rdf-syntax-ns#" ,
 "skos"="http://www.w3.org/2004/02/skos/core#" ,
 "prov"="http://www.w3.org/ns/prov#" ,
 "rdfs"="http://www.w3.org/2000/01/rdf-schema#" ,
 "dcat"="http://www.w3.org/ns/dcat#" ,
 "owl"= "http://www.w3.org/2002/07/owl#" ,
 "xsd"= "http://www.w3.org/2001/XMLSchema#" ,
 "qb"=  "http://purl.org/linked-data/cube#" ,
 "pav"= "http://purl.org/pav" ,
 "dct"= "http://purl.org/dc/terms/" ,
 "mms"= "http://rdf.cdisc.org/mms#" ,
 "cts"= "http://rdf.cdisc.org/ct/schema#",
 "rrdfqbcrnd0"="http://www.example.org/rrdfqbcrnd0/"
)  ;


```

When developing, the script is intented to run from the package root as
```{r, results='asis', eval=FALSE}
knit("inst/data-raw/create-qb-CDISC-prefix.Rmd")
```

knit runs the script in the data-raw directory, so it would be
expected to use pkg=".." to store the qbIClist in the data directory
However, it did not work - hence the setwd below-

Data are stored in the data directory, following [R packages by Hadley Wickham](http://r-pkgs.had.co.nz/data.html) and [Writing R Extensions](http://cran.r-project.org/doc/manuals/R-exts.pdf).


```{r, results='asis', eval=TRUE}

# not needed when working directory is package root
# but needed when run using the knitr command above
setwd("../..")
devtools::use_data(qbCDISCprefixes,overwrite=TRUE)

print("Done")

```
